---
title: "VS Code and Cursor on HPC"
author:
- name: MCCB Bioinformatics Core (Yaro Markov)
  affiliation: UMass Chan Medical School, Worcester, USA
output:
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
abstract: |
  A practical guide for running VS Code or Cursor directly on an HPC compute
  node through remote tunnels, plus recommended extensions for Python, R,
  notebooks, and agentic coding.
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  tidy = FALSE,
  eval = FALSE
)
```

# Overview

VS Code and Cursor can both connect to the HPC login node over SSH, but that is
not enough for interactive work that should run on a compute node. If you SSH to
the login node from the editor and then request an interactive session from the
editor terminal, only that terminal moves to the compute node. The editor server,
extensions, language servers, and coding agents remain on the login node.

To make the full editor session run on a compute node, start a remote tunnel from
inside an LSF compute job, then connect your local editor to that tunnel. The
commands below request a long interactive-style job for the tunnel. Adjust CPU,
memory, queue, and wall time as needed, especially for heavier analyses.

# Connect VS Code to a compute node

First SSH to the HPC login node from your local terminal:

```{bash}
ssh <username>@hpc.umassmed.edu
```

Submit an LSF job that logs in to VS Code tunnels and keeps the tunnel alive:

```{bash}
bsub -q long -n 4 -R rusage[mem=10G] -W 24:00 "echo '' | /share/bin/code tunnel user login --provider microsoft && while true; do /share/bin/code tunnel; sleep 5; done"
```

After submission, LSF prints a job ID. Watch the tunnel output with:

```{bash}
bpeek <job_id>
```

The output will show a browser link and a one-time code. Open the link locally,
enter the code, and authenticate with your Microsoft account.

On your local machine:

1. Open VS Code.
2. Install the **Remote - Tunnels** extension from the Extensions marketplace.
3. Open the command palette with `Command + Shift + P`.
4. Run **Remote-Tunnels: Connect to Tunnel**.
5. Choose **Microsoft account**.
6. Select the node name marked **Online**.

After the connection opens, the VS Code server, terminal, extensions, notebooks,
and agents are running on the compute node allocated by LSF.

# Connect Cursor to a compute node

Cursor uses almost the same workflow as VS Code, but requires a few setup steps
before the first tunnel session.

## Install the Remote Server extension in Cursor

Download the VSIX package for the Microsoft remote server extension:

<https://marketplace.visualstudio.com/_apis/public/gallery/publishers/ms-vscode/vsextensions/remote-server/latest/vspackage>

Then install it in Cursor:

1. Open the Cursor command palette with `Ctrl + Shift + P`.
2. Search for `VSIX`.
3. Choose **Extensions: Install from VSIX**.
4. Select the downloaded `.vsix` file.

## Install the Cursor CLI on HPC

SSH to the HPC login node from your local terminal:

```{bash}
ssh <username>@hpc.umassmed.edu
```

Download and unpack the Cursor CLI:

```{bash}
curl -Lk "https://api2.cursor.sh/updates/download-latest?os=cli-alpine-x64" --output cursor_cli.tar.gz
tar -xzf cursor_cli.tar.gz
```

Add the directory containing the `cursor` executable to your `PATH`. For example,
if the executable is in your home directory:

```{bash}
echo 'export PATH="$HOME:$PATH"' >> ~/.bashrc
source ~/.bashrc
```

Confirm that the CLI is available:

```{bash}
which cursor
cursor --help
```

## Start a Cursor tunnel job

Submit the tunnel job from the HPC login node:

```{bash}
bsub -q long -n 4 -R rusage[mem=10G] -W 24:00 "echo '' | ~/cursor tunnel user login --provider microsoft && while true; do ~/cursor tunnel; sleep 5; done"
```

Check the job output:

```{bash}
bpeek <job_id>
```

Open the displayed link locally, enter the code, and authenticate with your
Microsoft account. Then connect from Cursor using the same tunnel workflow as
VS Code: open the command palette, run **Connect to Tunnel**, choose your
Microsoft account, and select the compute node marked **Online**.

# Useful extensions

Install extensions separately in every place where you use the editor. A local
VS Code or Cursor install and a remote HPC tunnel session have separate extension
sets.

## Agentic coding

For VS Code, UMass Chan users may be eligible for free GitHub Copilot Pro through
GitHub Education. Follow GitHub's instructions for teachers and maintainers:

<https://docs.github.com/en/copilot/how-tos/copilot-on-github/set-up-copilot/enable-copilot/set-up-for-teachers-and-os-maintainers>

After authorization:

1. Install **GitHub Copilot Chat** from the VS Code marketplace.
2. Sign in with your GitHub account.
3. Use the Copilot Chat agent interface from the side panel.

You can also install the **Codex** extension to connect your ChatGPT account for
agentic coding.

Repository-specific agent instructions can help agents follow project standards:

<https://docs.github.com/en/copilot/how-tos/copilot-on-github/customize-copilot/add-custom-instructions/add-repository-instructions#creating-path-specific-custom-instructions>

## Python and Jupyter notebooks

Recommended extensions:

1. **Python Extension Pack**
2. **Pylance**
3. **Jupyter**

Register a project-specific Jupyter kernel from the active conda environment:

```{bash}
python -m ipykernel install \
  --user \
  --name "$PROJECT" \
  --display-name "Python (${PROJECT})"
```

## R interactive work

Recommended extension:

1. **R**

For an RStudio-like workflow, create a conda environment with R, language server
support, HTTP graphics, notebook support, and common analysis packages.

Create `envs/conda.yaml`:

```{bash}
mkdir -p envs
cat > envs/conda.yaml <<EOF
name: ${PROJECT}
channels:
  - conda-forge
  - bioconda
dependencies:
  - python=3.12
  - r-base
  - r-languageserver
  - r-httpgd
  - r-irkernel
  - r-jsonlite
  - r-httpuv
  - r-styler
  - r-lintr
  - jupyterlab
  - notebook
  - ipykernel
  - nbformat
  - jupyter_server
  - pandas
  - numpy
  - scipy
  - matplotlib
  - psutil
  - nodejs=20
EOF
```

Create and activate the environment:

```{bash}
mamba env create -f envs/conda.yaml
mamba activate "$PROJECT"
```

Optionally register an R kernel for Jupyter notebooks:

```{bash}
R -e 'IRkernel::installspec(name=Sys.getenv("CONDA_DEFAULT_ENV"), displayname=paste0("R (", Sys.getenv("CONDA_DEFAULT_ENV"), ")"))'
```

In VS Code, open the command palette, run **Preferences: Open User Settings
(JSON)**, and add:

```json
{
  "r.rpath.linux": "R",
  "r.alwaysUseActiveTerminal": true,
  "r.sessionWatcher": true
}
```

On HPC, add the following to `~/.Rprofile`:

```{r}
if (interactive() && Sys.getenv("RSTUDIO") == "") {
  source(file.path(
    Sys.getenv(if (.Platform$OS.type == "windows") "USERPROFILE" else "HOME"),
    ".vscode-R",
    "init.R"
  ))
}
```

With this configuration, VS Code launches R from the active terminal. That means
you can activate any conda environment first, then start R directly from the VS
Code terminal and use that environment's `R`. Otherwise, set `r.rpath.linux` in
the VS Code user settings to the exact path of the conda environment's R
executable.

To use R interactively in VS Code or Cursor:

1. Install the R extension in the editor session where you will work.
2. Activate your project conda environment and launch R.
3. Open an R script.
4. Open the R panel from the side bar.
5. Start HTTP graphics in the R session:

```{r}
library(httpgd)
hgd()
```

Select lines in the R script and press `Command + Enter` to run them. Variables
can be inspected from the R pane, and figures appear in separate panes.

## Extra extensions

These extensions are useful for common project work:

1. **Docker**
2. **EditorConfig**
3. **GitHub Pull Requests**
4. **Live Server**
5. **Live Share**
6. **R Debugger**
7. **Python Debugger**
8. **Python Environments**
9. **R Syntax**
10. **Rainbow CSV**
11. **Remote Development**
12. **YAML**
13. **SQLite3 Editor**

# Practical notes

Use tunnels only for active work, and stop the LSF job when you are finished so
the compute node is released. The tunnel job controls the compute resources
available to the editor session, so request enough memory, CPU, and wall time for
the analysis you plan to run.
